---
title: |
  \Large{Supplementary Materials: Development and Validation of the Other-Cause Comorbidity-Adjusted Mortality (OCCAM) Model for Clinical Use in Men with Prostate Cancer}
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(collapse=TRUE, fig.align='center', tidy=TRUE, tidy.opts=list(blank=TRUE, width.cutoff=40), warning=FALSE,message=FALSE)

rm(list=ls())

library(tidyverse)
library(gridExtra)
library(RColorBrewer)
library(tableone)
library(latex2exp)
library(labelled)
library(kableExtra)
library(survival)
library(survey)
library(randomForestSRC)
library(flexsurv)
library(pec)
library(timeROC)
library(cmprsk)

#We read in the final version of the NHANES data, saved at the end of OCM_modelbuilding.Rmd:
load("/Users/ecchase/Desktop/Research/Matt Prostate OCM/PCOtherCause/Data/nhanes_data_clean.RData")

#We read in the clean version of the PLCO data, saved at the end of OCM_modelvalidation.Rmd:
load("~/Box/PLCO Data/Prostate/plco_data_clean.RData")

cox_40 <-
  coxph(
    Surv(permth_exm, mortstat) ~ age_ctr_40 + diabetic + educ + hypertension + 
      marital2 + underweight + overweight2 + obese2 + smoker + stroke + 
      age_ctr_40 * diabetic + age_ctr_40 * educ + age_ctr_40 * hypertension + 
      age_ctr_40 * stroke + pc,
    data = mydata[mydata$inmodel4 == 1, ],
    x = TRUE,
    y = TRUE
  )
```

## Variable Definitions

We constructed some of our variables from other NHANES variables. In particular:

1.	We defined a patient as having hypertension if they reported a previous diagnosis of hypertension or if their blood pressure reading was hypertensive (we followed standard NHANES guidelines on constructing the blood pressure reading).

2.	We defined a patient as having diabetes if they reported a previous diagnosis of diabetes or if their blood glucose met clinical criteria for diabetes.

3.	We defined a patient as having high cholesterol if they reported high cholesterol, or if their blood work met clinical criteria for high cholesterol. 

4.	We defined patients as current smokers if they had smoked 100 cigarettes in their lifetime and reported smoking cigarettes every day or some days at present; we defined patients as former smokers if they had smoked 100 cigarettes in their lifetime and reported smoking not at all at present; we defined patients as never smokers if they had not smoked 100 cigarettes in their lifetime. 

5.	We defined being overweight as having BMI greater than or equal to 25 and less than 40 and being obese as having BMI greater than or equal to 40.

6.	We collapsed marital status into three categories: married if the patient was currently married or living with a partner, separated if the patient was widowed, divorced, or separated from their partner, and single if the patient had never married.

7.	We collapsed race into non-Hispanic white, non-Hispanic black, and all other races. 

For all other variables, we used unmodified results from the NHANES questionnaires about demographics, health insurance, health care access, mental health, and medical history. For more information on variable definition and construction, please see the repository posted at: https://github.com/blindedforreview/PCOtherCause. 

\newpage
## Supplementary Findings

\newpage
eFigure 1: Forest plot comparing the effect sizes of predictors from the unweighted and survey weighted versions of OCCAM fit in the NHANES training cohort of 7,369 men. The model also includes interactions between age and diabetes, education, hypertension, and stroke. 
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE}
table_model <- summary(cox_40)$conf.int
table_model[1, 1] <- exp(10*log(table_model[1, 1]))
table_model[1, 3] <- exp(10*log(table_model[1, 3]))
table_model[1, 4] <- exp(10*log(table_model[1, 4]))

nhanes_svy <-
  svydesign(
    data = mydata,
    id =  ~ psu,
    strata =  ~ strata,
    weights =  ~ sweights,
    nest = TRUE
  )
nhanes_sub4 <- subset(nhanes_svy, inmodel4 == 1)

cox_40_wt <-
  svycoxph(
    Surv(permth_exm, mortstat) ~ age_ctr_40 + diabetic + educ + hypertension + 
      marital2 + underweight + overweight2 + obese2 + smoker + stroke + 
      age_ctr_40 * diabetic + age_ctr_40 * educ + age_ctr_40 * hypertension + 
      age_ctr_40 * stroke + pc,
    design = nhanes_sub4)

table_model_wt <- summary(cox_40_wt)$conf.int
table_model_wt[1, 1] <- exp(10*log(table_model_wt[1, 1]))
table_model_wt[1, 3] <- exp(10*log(table_model_wt[1, 3]))
table_model_wt[1, 4] <- exp(10*log(table_model_wt[1, 4]))

cox_plotdata <- data.frame("Variable" = c("Age (per 10 years)", "Diabetes (ref: no diabetes)", "9th-11th Grade (ref: < 9th grade)", "High School (ref: < 9th grade)", "Some College (ref: < 9th grade)", "College (ref: < 9th grade)", "Hypertension (ref: no hypertension)", "Separated (ref: married)", "Single (ref: married)", "Current smoker (ref: never smoker)", "Former smoker (ref: never smoker)", "Stroke (ref: no stroke)", "Underweight, BMI <18.5 (ref: BMI 18.5-25)", "Overweight, BMI 25-40 (ref: BMI 18.5-25)", "Obese, BMI 40+ (ref: BMI 18.5-25)", "Prostate cancer (ref: no prostate cancer)"), "Effect" = c(table_model[1:16,1], table_model_wt[1:16,1]), "Lower" = c(table_model[1:16,3], table_model_wt[1:16,3]), "Upper" = c(table_model[1:16,4], table_model_wt[1:16,4]), "Label" = c(paste(round(table_model[1:16,1], digits=2), " (", round(table_model[1:16,3], digits=2), ",", round(table_model[1:16,4], digits=2), ")", sep=""), paste(round(table_model_wt[1:16,1], digits=2), " (", round(table_model_wt[1:16,3], digits=2), ",", round(table_model_wt[1:16,4], digits=2), ")", sep="")), "Model" = c(rep("Unweighted", 16), rep("Weighted", 16)))

cox_plotdata$Variable <- factor(cox_plotdata$Variable, levels= c("Age (per 10 years)", "Diabetes (ref: no diabetes)", "9th-11th Grade (ref: < 9th grade)", "High School (ref: < 9th grade)", "Some College (ref: < 9th grade)", "College (ref: < 9th grade)", "Hypertension (ref: no hypertension)", "Separated (ref: married)", "Single (ref: married)", "Current smoker (ref: never smoker)", "Former smoker (ref: never smoker)", "Stroke (ref: no stroke)", "Underweight, BMI <18.5 (ref: BMI 18.5-25)", "Overweight, BMI 25-40 (ref: BMI 18.5-25)", "Obese, BMI 40+ (ref: BMI 18.5-25)", "Prostate cancer (ref: no prostate cancer)"), ordered=TRUE)

adj_plot <- ggplot(cox_plotdata, aes(x = fct_rev(Variable), y = log(Effect), group = Model, color = Model)) + geom_point(position = position_dodge(width = 0.5)) + geom_errorbar(aes(ymin=log(Lower), ymax=log(Upper), width=0.1), position = position_dodge(width = 0.5)) + coord_flip() + labs(y = "Hazard Ratio and 95% CI", x = "", caption=TeX('Protective $\\leftarrow \\rightarrow$ Harmful' )) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme(panel.background = element_rect(fill="white"), panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.position="bottom", plot.caption = element_text(hjust = 0.3)) + scale_y_continuous(breaks=log(c(0.25, 0.5, 1, 2, 4)), labels=c("0.25", "0.5", "1.0", "2.0", "4.0")) + scale_color_manual(values = c("mediumpurple4", "gray74"))
 
adj_plot

ggsave("forest_plot_weighted.pdf", adj_plot)
rm(list=c("adj_plot", "cox_plotdata", "table_model"))
```

\newpage 

eFigure 2: Externally validated time-dependent area-under-the-curves (AUCs) at 5, 10, and 15 years of our other-cause comorbidity-adjusted mortality (OCCAM) model, the Social Security Administration 2001 actuarial life table predictions (SSA), and the National Vital Statistics System's 2001 life expectancy predictions (NVSS). Models were validated in the Prostate, Lung, Colon, and Ovarian Cancer Screening Trial (PLCO) cohort of 8,220 men with prostate cancer.
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE, cache = TRUE, eval = TRUE}
cox_age <-
  coxph(
    Surv(permth_exm, mortstat) ~ age_ctr_40 + pc,
    data = mydata[mydata$inmodel4 == 1, ],
    x = TRUE,
    y = TRUE
  )

times <- seq(60, 180, by=60)

plco_clean$permth_exm2 <- case_when(
  plco_clean$mortstat2==2 ~ 100000,
  plco_clean$mortstat2!= 2 ~ plco_clean$permth_exm
)

ssa <- read_csv("/Users/ecchase/Desktop/Research/Matt Prostate OCM/PCOtherCause/Data/ssa_2000.csv")
ssa$age <- seq(0, 119, by = 1)

nvss_black <- read_csv("/Users/ecchase/Desktop/Research/Matt Prostate OCM/PCOtherCause/Data/nvss_blackmales_2001.csv")
nvss_white <- read_csv("/Users/ecchase/Desktop/Research/Matt Prostate OCM/PCOtherCause/Data/nvss_whitemales_2001.csv")
nvss_other <- read_csv("/Users/ecchase/Desktop/Research/Matt Prostate OCM/PCOtherCause/Data/nvss_allmales_2001.csv")
nvss_other <- nvss_other[-102,]

nvss <- data.frame("Age" = seq(0, 100, by = 1), "White" = nvss_white$Expectancy, "Black" = nvss_black$Expectancy, "Other" = nvss_other$Expectancy)

plco_clean$ssa <- NA
plco_clean$nvss <- NA
for (i in 1:nrow(plco_clean)){
  plco_clean$ssa[i] <- ssa$Expectancy[which(ssa$age==plco_clean$age[i])]
  if (plco_clean$race2[i]=="NHW"){
    plco_clean$nvss[i] <- nvss$White[which(nvss$Age==plco_clean$age[i])]
  } else if (plco_clean$race2[i]=="NHB"){
    plco_clean$nvss[i] <- nvss$Black[which(nvss$Age==plco_clean$age[i])]
  } else if (plco_clean$race2[i]=="Other"){
    plco_clean$nvss[i] <- nvss$Other[which(nvss$Age==plco_clean$age[i])]
  }
}
```

```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE, cache = TRUE, eval = TRUE}
#cox40_plco <- pec::cindex(object=cox_40, formula = Surv(permth_exm, mortstat) ~ age_ctr_40 + diabetic + educ + hypertension + marital2 + underweight + overweight2 + obese2 + smoker + stroke + age_ctr_40*diabetic + age_ctr_40*educ + age_ctr_40*hypertension + age_ctr_40*stroke + pc, data=plco_clean, eval.times = times)

cox40predict <- predict(object=cox_40, newdata = plco_clean, type="lp")
coxagepredict <- predict(object=cox_age, newdata = plco_clean, type="lp")

cox40roc_cs <-
  timeROC(
    T = plco_clean$permth_exm,
    delta = plco_clean$mortstat,
    marker = cox40predict,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

cox40roc_cs_ci <- confint(cox40roc_cs)

coxageroc_cs <-
  timeROC(
    T = plco_clean$permth_exm,
    delta = plco_clean$mortstat,
    marker = coxagepredict,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

coxageroc_cs_ci <- confint(coxageroc_cs)

cox40roc_cr <-
  timeROC(
    T = plco_clean$permth_exm2,
    delta = plco_clean$mortstat,
    marker = cox40predict,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

cox40roc_cr_ci <- confint(cox40roc_cr)

coxageroc_cr <-
  timeROC(
    T = plco_clean$permth_exm2,
    delta = plco_clean$mortstat,
    marker = coxagepredict,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

coxageroc_cr_ci <- confint(coxageroc_cr)

ssa_roc_cs <-
  timeROC(
    T = plco_clean$permth_exm,
    delta = plco_clean$mortstat,
    marker = 1/plco_clean$ssa,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

ssa_cs_ci <- confint(ssa_roc_cs)

ssa_roc_cr <-
  timeROC(
    T = plco_clean$permth_exm2,
    delta = plco_clean$mortstat,
    marker = 1/plco_clean$ssa,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

ssa_cr_ci <- confint(ssa_roc_cr)

nvss_roc_cs <-
  timeROC(
    T = plco_clean$permth_exm,
    delta = plco_clean$mortstat,
    marker = 1/plco_clean$nvss,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

nvss_roc_cs_ci <- confint(nvss_roc_cs)

nvss_roc_cr <-
  timeROC(
    T = plco_clean$permth_exm2,
    delta = plco_clean$mortstat,
    marker = 1/plco_clean$nvss,
    cause = 1,
    weighting = "marginal",
    times = times,
    iid = TRUE
  )

nvss_roc_cr_ci <- confint(nvss_roc_cr)

subperf <- data.frame("Model" = rep(c("OCCAM", "Age Model", "SSA", "NVSS"), 6), "Metric" = c(rep("CS", 12), rep("CR", 12)), "Year" = c(rep(5, 4), rep(10, 4), rep(15, 4)), "AUC" = c(cox40roc_cs$AUC[1], coxageroc_cs$AUC[1], ssa_roc_cs$AUC[1], nvss_roc_cs$AUC[1], cox40roc_cs$AUC[2], coxageroc_cs$AUC[2], ssa_roc_cs$AUC[2], nvss_roc_cs$AUC[2], cox40roc_cs$AUC[3], coxageroc_cs$AUC[3], ssa_roc_cs$AUC[3], nvss_roc_cs$AUC[3], cox40roc_cr$AUC[1], coxageroc_cr$AUC[1], ssa_roc_cr$AUC[1], nvss_roc_cr$AUC[1], cox40roc_cr$AUC[2], coxageroc_cr$AUC[2], ssa_roc_cr$AUC[2], nvss_roc_cr$AUC[2], cox40roc_cr$AUC[3], coxageroc_cr$AUC[3], ssa_roc_cr$AUC[3], nvss_roc_cr$AUC[3]), "Lower" = c(cox40roc_cs_ci$CB_AUC[1,1], coxageroc_cs_ci$CB_AUC[1,1], ssa_cs_ci$CB_AUC[1,1], nvss_roc_cs_ci$CB_AUC[1,1], cox40roc_cs_ci$CB_AUC[2,1], coxageroc_cs_ci$CB_AUC[2,1], ssa_cs_ci$CB_AUC[2,1], nvss_roc_cs_ci$CB_AUC[2,1], cox40roc_cs_ci$CB_AUC[3,1], coxageroc_cs_ci$CB_AUC[3,1], ssa_cs_ci$CB_AUC[3,1], nvss_roc_cs_ci$CB_AUC[3,1]), "Upper" = c(cox40roc_cs_ci$CB_AUC[1,2], coxageroc_cs_ci$CB_AUC[1,2], ssa_cs_ci$CB_AUC[1,2], nvss_roc_cs_ci$CB_AUC[1,2], cox40roc_cs_ci$CB_AUC[2,2], coxageroc_cs_ci$CB_AUC[2,2], ssa_cs_ci$CB_AUC[2,2], nvss_roc_cs_ci$CB_AUC[2,2], cox40roc_cs_ci$CB_AUC[3,2], coxageroc_cs_ci$CB_AUC[3,2], ssa_cs_ci$CB_AUC[3,2], nvss_roc_cs_ci$CB_AUC[3,2]))
                      
#aucplot <- ggplot(data = subperf, aes (x = Year, y = AUC, group = Model, color = Model)) + geom_point() + geom_line() + geom_errorbar(aes(ymin = Lower/100, ymax = Upper/100, width = 0.05)) + facet_grid(~Metric) + ylab("IPCW Time-Dependent AUC") + xlab("Time (Years)") + scale_x_continuous(breaks = c(0, 5, 10, 15)) + theme_bw() + scale_color_manual(values = c("mediumpurple4", "darkblue", "gray74", "darkred"))
```

```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE, eval = TRUE}
subperf$Model <- factor(subperf$Model, levels=c("OCCAM", "NVSS", "SSA", "Age Model"), ordered = TRUE)

aucplot <- ggplot(data = subperf[subperf$Metric=="CS" & subperf$Model != "Age Model",], aes (x = Year, y = AUC, group = Model, color = Model)) + geom_point(position = position_dodge(width = 0.4)) + geom_line() + geom_errorbar(aes(ymin = Lower/100, ymax = Upper/100, width = 0.05), position = position_dodge(width = 0.4)) + ylab("IPCW Time-Dependent AUC") + xlab("Time (Years)") + scale_x_continuous(breaks = c(0, 5, 10, 15)) + theme_bw() + scale_color_manual(values = c("darkred", "mediumpurple4", "grey74")) + scale_y_continuous(breaks = c(0.5, 0.75, 1)) + coord_cartesian(ylim = c(0.5, 1))

aucplot

ggsave("auc_performance.pdf", aucplot)

rm(list=c("cox40roc_cs", "cox40roc_cr", "coxageroc_cs", "coxageroc_cr", "ssa_roc_cs", "ssa_roc_cr", "nvss_roc_cs", "nvss_roc_cr", "subperf", "cox40roc_cs_ci", "cox40roc_cr_ci", "coxageroc_cs_ci", "coxageroc_cr_ci", "ssa_roc_cs_ci", "ssa_roc_cr_ci", "nvss_roc_cs_ci", "nvss_roc_cr_ci", "aucplot"))
```

\newpage

eFigure 3: Calibration performance of other-cause comorbidity-adjusted mortality (OCCAM) model and the Social Security Administration's 2001 actuarial life table predictions (SSA) in the Prostate, Lung, Colon, and Ovarian Cancer Screening Trial (PLCO) cohort of 8,220 men with prostate cancer.
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE}
km_predictions <- survfit(cox_40, plco_clean)
cox_40_calib <- coxph(Surv(permth_exm, mortstat) ~ linpred, data = plco_clean)
km_predictions_calib <- survfit(cox_40_calib, plco_clean)
myfunc <- function(x, time){
  inds <- which(x-0.5 <= 0)
  if (identical(inds, integer(0))){
    mytime <- 202
  } else{
    mytime <- time[min(inds)]
  }
  return(mytime)
}
medsurv <- apply(X = km_predictions$surv, MARGIN = 2, FUN = myfunc, time = km_predictions$time)
medsurv_calib <- apply(X = km_predictions_calib$surv, MARGIN = 2, FUN = myfunc, time = km_predictions_calib$time)
medsurv2 <- data.frame(medsurv)
medsurv2_calib <- data.frame(medsurv_calib)

plco_clean$medsurv <- medsurv2$medsurv
plco_clean$medsurv_calib <- medsurv2_calib$medsurv_calib

cox_group <- case_when(
  plco_clean$medsurv < 60 ~ 2.5,
  plco_clean$medsurv >= 60 & plco_clean$medsurv < 84 ~ 6,
  plco_clean$medsurv >= 84 & plco_clean$medsurv < 120 ~ 8.5,
  plco_clean$medsurv >= 120 & plco_clean$medsurv < 144 ~ 11,
  plco_clean$medsurv >= 144 & plco_clean$medsurv < 180 ~ 13.5,
  plco_clean$medsurv >= 180 ~ 15
)

cox_group_calib <- case_when(
  plco_clean$medsurv_calib < 60 ~ 2.5,
  plco_clean$medsurv_calib >= 60 & plco_clean$medsurv_calib < 84 ~ 6,
  plco_clean$medsurv_calib >= 84 & plco_clean$medsurv_calib < 120 ~ 8.5,
  plco_clean$medsurv_calib >= 120 & plco_clean$medsurv_calib < 144 ~ 11,
  plco_clean$medsurv_calib >= 144 & plco_clean$medsurv_calib < 180 ~ 13.5,
  plco_clean$medsurv_calib >= 180 ~ 15
)

ssa_group <- case_when(
  plco_clean$ssa < 5 ~ 2.5,
  plco_clean$ssa >= 5 & plco_clean$ssa < 7 ~ 6,
  plco_clean$ssa >= 7 & plco_clean$ssa < 10 ~ 8.5,
  plco_clean$ssa >= 10 & plco_clean$ssa < 12 ~ 11,
  plco_clean$ssa >= 12 & plco_clean$ssa < 15 ~ 13.5,
  plco_clean$ssa >= 15 ~ 15
)

cox_5 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==2.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==2.5,])$time))
coxcalib_5 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==2.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==2.5,])$time))
ssa_5 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==2.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==2.5,])$time))
cox_7 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==6,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==6,])$time))
coxcalib_7 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==6,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==6,])$time))
ssa_7 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==6,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==6,])$time))
cox_10 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==8.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==8.5,])$time))
coxcalib_10 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==8.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==8.5,])$time))
ssa_10 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==8.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==8.5,])$time))
cox_12 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==11,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==11,])$time))
coxcalib_12 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==11,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==11,])$time))
ssa_12 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==11,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==11,])$time))
cox_15 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==13.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==13.5,])$time))
coxcalib_15 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==13.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==13.5,])$time))
ssa_15 <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==13.5,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==13.5,])$time))
cox_15p <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==15,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==15,])$time))
coxcalib_15p <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==15,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==15,])$time))
ssa_15p <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==15,])$surv, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==15,])$time))

cox_5l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==2.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==2.5,])$time))
coxcalib_5l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==2.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==2.5,])$time))
ssa_5l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==2.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==2.5,])$time))
cox_7l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==6,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==6,])$time))
coxcalib_7l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==6,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==6,])$time))
ssa_7l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==6,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==6,])$time))
cox_10l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==8.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==8.5,])$time))
coxcalib_10l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==8.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==8.5,])$time))
ssa_10l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==8.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==8.5,])$time))
cox_12l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==11,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==11,])$time))
coxcalib_12l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==11,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==11,])$time))
ssa_12l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==11,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==11,])$time))
cox_15l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==13.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==13.5,])$time))
coxcalib_15l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==13.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==13.5,])$time))
ssa_15l <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==13.5,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==13.5,])$time))
cox_15pl <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==15,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==15,])$time))
coxcalib_15pl <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==15,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==15,])$time))
ssa_15pl <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==15,])$lower, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==15,])$time))

cox_5u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==2.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==2.5,])$time))
coxcalib_5u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==2.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==2.5,])$time))
ssa_5u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==2.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==2.5,])$time))
cox_7u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==6,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==6,])$time))
coxcalib_7u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==6,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==6,])$time))
ssa_7u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==6,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==6,])$time))
cox_10u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==8.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==8.5,])$time))
coxcalib_10u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==8.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==8.5,])$time))
ssa_10u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==8.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==8.5,])$time))
cox_12u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==11,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==11,])$time))
coxcalib_12u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==11,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==11,])$time))
ssa_12u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==11,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==11,])$time))
cox_15u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==13.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==13.5,])$time))
coxcalib_15u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==13.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==13.5,])$time))
ssa_15u <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==13.5,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==13.5,])$time))
cox_15pu <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==15,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group==15,])$time))
coxcalib_15pu <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==15,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[cox_group_calib==15,])$time))
ssa_15pu <- data.frame(myfunc(x = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==15,])$upper, time = survfit(Surv(permth_exm, mortstat) ~ 1, data = plco_clean[ssa_group==15,])$time))

calibration_data <- data.frame("Time" = rep(c(2.5, 6, 8.5, 11, 13.5, 15), 3), "Model" = c(rep("NHANES-Calibrated OCCAM", 6), rep("PLCO-Calibrated OCCAM", 6), rep("SSA", 6)), "Observed" = c(cox_5$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_7$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_10$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_12$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_15$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_15p$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., coxcalib_5$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_7$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_10$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_12$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_15$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_15p$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib....,ssa_5$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_7$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_10$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_12$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_15$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_15p$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group....)/12, "Lower" = c(cox_5l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_7l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_10l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_12l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_15l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_15pl$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., coxcalib_5l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_7l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_10l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_12l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_15l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_15pl$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib....,ssa_5l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_7l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_10l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_12l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_15l$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_15pl$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group....)/12, "Upper" = c(cox_5u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_7u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_10u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_12u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_15u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., cox_15pu$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group...., coxcalib_5u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_7u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_10u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_12u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_15u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib...., coxcalib_15pu$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.cox_group_calib....,ssa_5u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_7u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_10u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_12u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_15u$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group...., ssa_15pu$myfunc.x...survfit.Surv.permth_exm..mortstat....1..data...plco_clean.ssa_group....)/12)

calibration_data[calibration_data==202/12] <- 25

#myplot <- ggplot() + geom_line(data = calibration_data, aes(x=Time, y=Observed, 
                  #                                              group = Model, 
                #                                                color = Model)) +
 # geom_errorbar(data = calibration_data, aes(x = Time, ymin = Lower, ymax = Upper, group = Model, color = Model, width = 0.5)) + 
#  scale_x_continuous(name = "Predicted Life Expectancy", 
#                     breaks = seq(2, 16, by=3)) + 
#  scale_y_continuous(name = "Observed Life Expectancy", 
#                     breaks = seq(2, 23, by=3)) + theme_bw() + 
 # geom_abline(slope = 1, intercept = 0) + scale_color_manual(values = c("darkred", "mediumpurple4", "grey74"), name = "Model") + theme(legend.position = "bottom") + coord_cartesian(ylim = c(2, 23), xlim = c(2,16))

myplot <- ggplot() + geom_line(data = calibration_data, aes(x=Time, y=Observed, 
                                                                group = Model, 
                                                                linetype = Model)) +
  geom_errorbar(data = calibration_data, aes(x = Time, ymin = Lower, ymax = Upper, group = Model, linetype = Model, width = 0.5)) + 
  scale_x_continuous(name = "Predicted Life Expectancy", 
                     breaks = seq(2, 16, by=3)) + 
  scale_y_continuous(name = "Observed Life Expectancy", 
                     breaks = seq(2, 23, by=3)) + theme_bw() + 
  geom_abline(slope = 1, intercept = 0) + theme(legend.position = "bottom") + coord_cartesian(ylim = c(2, 23), xlim = c(2,16)) + scale_linetype_manual(values = c("twodash", "longdash", "dotted"))

myplot

ggsave("calibration_plot.pdf", myplot)

rm(list=c("myplot", "ssa_5", "ssa_5l", "ssa_5u", "ssa_7", "ssa_7l", "ssa_7u", "ssa_10", "ssa_10l", "ssa_10u", "ssa_12", "ssa_12u", "ssa_12l", "ssa_15", "ssa_15u", "ssa_15l", "ssa_15p", "ssa_15pl", "ssa_15pu", "ssa_5", "ssa_5l", "ssa_5u", "ssa_7", "ssa_7l", "ssa_7u", "ssa_10", "ssa_10l", "ssa_10u", "ssa_12", "ssa_12u", "ssa_12l", "ssa_15", "ssa_15u", "ssa_15l", "ssa_15p", "ssa_15pl", "ssa_15pu"))
```

eTable 1: Cause-specific and competing-risks time-dependent AUCs of OCCAM in the Prostate, Lung, Colon, and Ovarian Cancer Screening Trial (PLCO) validation cohort of 8,220 men.
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE}
#And now we obtain estimates of the C-index and time-dependent AUC from our model in the main treatment groups: prostatectomy, radiation, and radiation + hormone:
times <- seq(6, 180, by=6)

roc_css <- timeROC(T = plco_clean$permth_exm, delta = plco_clean$mortstat, marker = plco_clean$linpred, cause = 1, weighting="marginal", times = times)

plco_clean$permth_exm2 <- case_when(
  plco_clean$mortstat2==2 ~ 100000,
  plco_clean$mortstat2!= 2 ~ plco_clean$permth_exm
)

roc_cr <- timeROC(T = plco_clean$permth_exm2, delta = plco_clean$mortstat, marker = plco_clean$linpred, cause = 1, weighting="marginal", times = times)

subperf_trt <- data.frame("Metric" = c("Cause-Specific AUC", "Competing-Risks AUC"), "Year5" = round(c(roc_css$AUC[10], roc_cr$AUC[10]), digits=2), "Year10" = round(c(roc_css$AUC[20], roc_cr$AUC[20]), digits=2), "Year15" = round(c(roc_css$AUC[30], roc_cr$AUC[30]), digits=2))

rownames(subperf_trt) <- NULL
kable(subperf_trt, col.names = c("Metric", "5 Years", "10 Years", "15 Years")) %>% column_spec(column = 1, width = "3cm") %>% column_spec(column = 2:4, width = "1.75cm")
```

\newpage
eTable 2: Time-dependent AUC of OCCAM in the three main treatment groups in the Prostate, Lung, Colon, and Ovarian Cancer Screening Trial (PLCO) validation cohort of 8,220 men.
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE}
#And now we obtain estimates of the C-index and time-dependent AUC from our model in the main treatment groups: prostatectomy, radiation, and radiation + hormone:
times <- seq(6, 180, by=6)

cox40predict_prostatectomy <- predict(object=cox_40, newdata = plco_clean[plco_clean$primary_tx=="Prostatectomy",], type="lp")

cox40predict_radiation <- predict(object=cox_40, newdata = plco_clean[plco_clean$primary_tx=="Radiation alone",], type="lp")

cox40predict_rtadt <- predict(object=cox_40, newdata = plco_clean[plco_clean$primary_tx=="Radiation + hormone",], type="lp")

prostatectomy_roc_css <- timeROC(T = plco_clean$permth_exm[plco_clean$primary_tx=="Prostatectomy"], delta = plco_clean$mortstat[plco_clean$primary_tx=="Prostatectomy"], marker = cox40predict_prostatectomy, cause = 1, weighting="marginal", times = times)

radiation_roc_css <- timeROC(T = plco_clean$permth_exm[plco_clean$primary_tx=="Radiation alone"], delta = plco_clean$mortstat[plco_clean$primary_tx=="Radiation alone"], marker = cox40predict_radiation, cause = 1, weighting="marginal", times = times)

rtadt_roc_css <- timeROC(T = plco_clean$permth_exm[plco_clean$primary_tx=="Radiation + hormone"], delta = plco_clean$mortstat[plco_clean$primary_tx=="Radiation + hormone"], marker = cox40predict_rtadt, cause = 1, weighting="marginal", times = times)

plco_clean$permth_exm2 <- case_when(
  plco_clean$mortstat2==2 ~ 100000,
  plco_clean$mortstat2!= 2 ~ plco_clean$permth_exm
)

prostatectomy_roc_cr <- timeROC(T = plco_clean$permth_exm2[plco_clean$primary_tx=="Prostatectomy"], delta = plco_clean$mortstat[plco_clean$primary_tx=="Prostatectomy"], marker = cox40predict_prostatectomy, cause = 1, weighting="marginal", times = times)

radiation_roc_cr <- timeROC(T = plco_clean$permth_exm2[plco_clean$primary_tx=="Radiation alone"], delta = plco_clean$mortstat[plco_clean$primary_tx=="Radiation alone"], marker = cox40predict_radiation, cause = 1, weighting="marginal", times = times)

rtadt_roc_cr <- timeROC(T = plco_clean$permth_exm2[plco_clean$primary_tx=="Radiation + hormone"], delta = plco_clean$mortstat[plco_clean$primary_tx=="Radiation + hormone"], marker = cox40predict_rtadt, cause = 1, weighting="marginal", times = times)

subperf_trt <- data.frame("Treatment" = c("Prostatectomy", "Radiation alone", "Radiation + hormone"), "Year5_AUC_CSS" = round(c(prostatectomy_roc_css$AUC[10], radiation_roc_css$AUC[10], rtadt_roc_css$AUC[10]), digits=2), "Year10_AUC_CSS" = round(c(prostatectomy_roc_css$AUC[20], radiation_roc_css$AUC[20],rtadt_roc_css$AUC[20]), digits=2), "Year15_AUC_CSS" = round(c(prostatectomy_roc_css$AUC[30], radiation_roc_css$AUC[30], rtadt_roc_css$AUC[30]), digits=2), "Year5_AUC_CR" = round(c(prostatectomy_roc_cr$AUC[10], radiation_roc_cr$AUC[10], rtadt_roc_cr$AUC[10]), digits=2), "Year10_AUC_CR" = round(c(prostatectomy_roc_cr$AUC[20], radiation_roc_cr$AUC[20],rtadt_roc_cr$AUC[20]), digits=2), "Year15_AUC_CR" = round(c(prostatectomy_roc_cr$AUC[30], radiation_roc_cr$AUC[30], rtadt_roc_cr$AUC[30]), digits=2))

rownames(subperf_trt) <- NULL
kable(subperf_trt, col.names = c("Treatment", "5 Years", "10 Years", "15 Years", "5 Years", "10 Years", "15 Years")) %>%
add_header_above(c(" ", "Cause-Specific AUC" = 3, "Competing-Risks AUC" = 3), escape=FALSE) %>% column_spec(column = 1, width = "3cm") %>% column_spec(column = 2:7, width = "1.75cm")

#subperf_trt <- subperf_trt[,1:4]
#kable(subperf_trt, col.names = c("Treatment", "5 Years", "10 Years", "15 Years")) %>%
#column_spec(column = 1, width = "3cm") %>% column_spec(column = 2:7, width = "1.75cm")

rm(list=c("subperf_trt", "cox40predict_prostatectomy", "cox40predict_radiation", "cox40predict_rtadt", "radiation_roc", "prostatectomy_roc", "rtadt_roc", "cox40_radiation", "cox40_rtadt", "cox40_prostatectomy"))
```

\newpage
eTable 3: Characteristics of a cohort of 7,596 men from the Prostate, Lung, Colorectal, and Ovarian Cancer Screening Trial, grouped by OCCAM-predicted median survival time. 
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE}
#Figure 6: Treatment concordance with OCM predictions:
prostatedata <- read_csv("~/Box/PLCO Data/Prostate/pros_data_nov18_d070819.csv.zip")
cancer <- filter(prostatedata, pros_cancer == 1)
rm(list = c("prostatedata"))

#We filter to the list of variables we need for validation:
plco <-
  select(
    cancer,
    race7,
    pros_exitage,
    diabetes_f,
    hyperten_f,
    stroke_f,
    arthrit_f,
    bronchit_f,
    emphys_f,
    hearta_f,
    liver_comorbidity,
    educat,
    marital,
    cig_stat,
    bmi_curr,
    is_dead_with_cod,
    mortality_exitdays,
    pros_exitdays,
    f_dthp,
    primary_trtp,
    pros_stage_n, 
    pros_stage_t, 
    pros_gleason, 
    dx_psa
  )

plco$nstage <- case_when(
  plco$pros_stage_n==0 ~ "N0",
  plco$pros_stage_n==100 ~ "N1",
  plco$pros_stage_n==200 ~ "N2",
  plco$pros_stage_n==999 ~ "NX"
)

plco$tstage <- case_when(
  plco$pros_stage_t==0 ~ "T0",
  plco$pros_stage_t==100 | plco$pros_stage_t==110 | plco$pros_stage_t==120 | plco$pros_stage_t==130 | plco$pros_stage_t==210 ~ "T1-T2a",
  plco$pros_stage_t==200 | plco$pros_stage_t==220 | plco$pros_stage_t==230 ~ "T2b-T2c",
  plco$pros_stage_t==300 | plco$pros_stage_t==310 | plco$pros_stage_t==320 | plco$pros_stage_t==330 | plco$pros_stage_t==400 | plco$pros_stage_t==410 | plco$pros_stage_t==420 ~ "T3a-T4",
  plco$pros_stage_t==999 ~ "TX"
)

plco$gleason <- case_when(
  plco$pros_gleason <= 6 ~ "<= 6",
  plco$pros_gleason == 7 ~ "7",
  plco$pros_gleason ==8 ~ "8",
  plco$pros_gleason ==9 ~ "9",
  plco$pros_gleason == 10 ~ "10",
  plco$pros_gleason == 99 ~ NA_character_
)

#We define derived variables:
plco$race2 <- case_when(
  plco$race7 == 1 ~ "NHW",
  plco$race7 == 2 ~ "NHB",
  plco$race7 >= 3 & plco$race7 <= 6 ~ "Other",
  plco$race7 == 7 ~ NA_character_
)

plco$educ <- case_when(
  plco$educat == 1 ~ "Less than 9th grade",
  plco$educat == 2 ~ "9th-11th grade",
  plco$educat == 3 ~ "HS graduate",
  plco$educat == 4 | plco$educat == 5 ~ "Some college",
  plco$educat == 6 | plco$educat == 7 ~ "College graduate",
  is.na(plco$educat) ~ NA_character_
)

plco$marital2 <- case_when(
  plco$marital == 1 ~ "Married",
  plco$marital == 2 |
    plco$marital == 3 | plco$marital == 4 ~ "Separated",
  plco$marital == 5 ~ "Single",
  is.na(plco$marital) ~ NA_character_
)

plco$smoker <- case_when(plco$cig_stat == 0 ~ "Never",
                         plco$cig_stat == 1 ~ "Current",
                         plco$cig_stat == 2 ~ "Former")

plco$primary_tx <- case_when(
  plco$primary_trtp == 0 ~ "No cancer",
  plco$primary_trtp == 1 ~ "Prostatectomy",
  plco$primary_trtp == 2 ~ "Radiation alone",
  plco$primary_trtp == 3 ~ "Radiation + hormone",
  plco$primary_trtp == 4 ~ "Hormone alone",
  plco$primary_trtp == 5 ~ "Other ablative treatment",
  plco$primary_trtp == 6 ~ "No known treatment",
  plco$primary_trtp == 10 ~ NA_character_
)

plco$underweight <-
  ifelse(is.na(plco$bmi_curr), NA, ifelse(plco$bmi_curr < 18.5,
                                          1, 0))
plco$overweight2 <-
  ifelse(is.na(plco$bmi_curr),
         NA,
         ifelse(plco$bmi_curr >= 25
                &
                  plco$bmi_curr < 40,
                1, 0))
plco$obese2 <-
  ifelse(is.na(plco$bmi_curr), NA, ifelse(plco$bmi_curr >= 40, 1, 0))
plco$overweight_ex <-
  ifelse(is.na(plco$bmi_curr),
         NA,
         ifelse(plco$bmi_curr >= 25
                &
                  plco$bmi_curr < 30,
                1, 0))
plco$obese <-
  ifelse(is.na(plco$bmi_curr), NA, ifelse(plco$bmi_curr >= 30, 1, 0))

plco$mortstat <- case_when(
  plco$is_dead_with_cod == 1 & plco$f_dthp == 0 ~ 1,
  plco$is_dead_with_cod == 1 & plco$f_dthp == 1 ~ 0,
  plco$is_dead_with_cod == 0 ~ 0
)

plco$mortstat2 <- case_when(
  plco$is_dead_with_cod == 1 & plco$f_dthp == 0 ~ 1,
  plco$is_dead_with_cod == 1 & plco$f_dthp == 1 ~ 2,
  plco$is_dead_with_cod == 0 ~ 0
)

plco$permth_exm <-
  round((plco$mortality_exitdays - plco$pros_exitdays) / 30)

#We select variables and do necessary renaming
plco_clean2 <- mutate(
  plco,
  age = pros_exitage,
  diabetic = diabetes_f,
  hypertension = hyperten_f,
  stroke = stroke_f,
  arthritis = arthrit_f,
  bronch = bronchit_f,
  emphysema = emphys_f,
  mi_chd = hearta_f,
  liver = liver_comorbidity,
  pc = 1
) %>%
  select(
    race2,
    age,
    diabetic,
    hypertension,
    stroke,
    arthritis,
    bronch,
    emphysema,
    mi_chd,
    liver,
    educ,
    marital2,
    smoker,
    underweight,
    overweight2,
    obese2,
    overweight_ex,
    obese,
    pc,
    mortstat,
    mortstat2,
    permth_exm,
    primary_tx,
    nstage,
    tstage,
    gleason,
    dx_psa
  )

#We define factors:
plco_clean2$race2 <- factor(plco_clean2$race2)
plco_clean2$educ <- factor(plco_clean2$educ, levels = c("Less than 9th grade", "9th-11th grade", "HS graduate", "Some college", "College graduate"), ordered=TRUE)
plco_clean2$marital2 <- factor(plco_clean2$marital2)
plco_clean2$smoker <- factor(plco_clean2$smoker)
plco_clean2$pc <-
  factor(plco_clean2$pc,
         levels = c(0, 1),
         labels = c("No PC", "PC"))
plco_clean2$hypertension <-
  factor(plco_clean2$hypertension,
         levels = c(0, 1),
         labels = c("No", "Yes"))
plco_clean2$diabetic <- factor(plco_clean2$diabetic,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))
plco_clean2$stroke <- factor(plco_clean2$stroke,
                            levels = c(0, 1),
                            labels = c("No", "Yes"))
plco_clean2$arthritis <-
  factor(plco_clean2$arthritis,
         levels = c(0, 1),
         labels = c("No", "Yes"))
plco_clean2$bronch <- factor(plco_clean2$bronch,
                            levels = c(0, 1),
                            labels = c("No", "Yes"))
plco_clean2$emphysema <-
  factor(plco_clean2$emphysema,
         levels = c(0, 1),
         labels = c("No", "Yes"))
plco_clean2$mi_chd <- factor(plco_clean2$mi_chd,
                            levels = c(0, 1),
                            labels = c("No", "Yes"))
plco_clean2$liver <- factor(plco_clean2$liver,
                           levels = c(0, 1),
                           labels = c("No", "Yes"))
plco_clean2$underweight <-
  factor(plco_clean2$underweight,
         levels = c(0, 1),
         labels = c("No", "Yes"))
plco_clean2$overweight2 <-
  factor(plco_clean2$overweight2,
         levels = c(0, 1),
         labels = c("No", "Yes"))
plco_clean2$obese2 <- factor(plco_clean2$obese2,
                            levels = c(0, 1),
                            labels = c("No", "Yes"))

plco_clean2$tstage <- factor(plco_clean2$tstage)
plco_clean2$nstage <- factor(plco_clean2$nstage)
plco_clean2$gleason <- factor(plco_clean2$gleason, levels=c("<= 6","7","8","9","10"), ordered=TRUE)

plco_clean2 <- na.omit(plco_clean2)

plco_clean2$age_ctr_40 <- plco_clean2$age - 60.34387 #scaling the age 40+ cohort

km_predictions <- survfit(cox_40, plco_clean2)
myfunc <- function(x){
  inds <- which(x-0.5 <= 0)
  if (identical(inds, integer(0))){
    mytime <- 202
  } else{
    mytime <- min(inds)
  }
  return(mytime)
}

medsurv <- apply(X = km_predictions$surv, MARGIN = 2, FUN = myfunc)
medsurv2 <- data.frame(medsurv)

causespecific_group <- case_when(
  medsurv2$medsurv >= 180 ~ "15+",
  medsurv2$medsurv >= 120 & medsurv2$medsurv < 180 ~ "10-15",
  medsurv2$medsurv >= 60 & medsurv2$medsurv < 120 ~ "5-10",
  medsurv2$medsurv < 60 ~ "< 5"
)

plco_clean2$causespecific_group <- factor(causespecific_group, levels=c("15+", "10-15", "5-10", "< 5"), ordered=TRUE)

plco_clean2$lifeexpectancy <- medsurv2$medsurv

plco_clean2$weight <- case_when(
  plco_clean2$underweight=="Yes"~ "Underweight (BMI < 18.5)",
  plco_clean2$underweight=="No" & plco_clean2$overweight2=="No" & plco_clean2$obese2=="No"~ "Normal weight (BMI 18.5-25)",
  plco_clean2$overweight2=="Yes" ~ "Overweight (BMI 25-40)",
  plco_clean2$obese2=="Yes" ~ "Obese (BMI 40+)"
)

plco_clean2$weight <- factor(plco_clean2$weight, levels=c("Underweight (BMI < 18.5)", "Normal weight (BMI 18.5-25)", "Overweight (BMI 25-40)", "Obese (BMI 40+)"), ordered=TRUE)

plco_clean2$mortstat <- factor(plco_clean2$mortstat2, levels=c(0, 1, 2), labels = c("Alive", "Died of OC", "Died of PC"))

plco_clean2$pc_risk_group <- case_when(
  plco_clean2$tstage=="T1-T2a" & plco_clean2$gleason =="<= 6" & plco_clean2$dx_psa < 10 & plco_clean2$nstage=="N0" ~ "Low",
  plco_clean2$tstage!= "T3a-T4" & plco_clean2$gleason == "7" & plco_clean2$dx_psa <= 20 & plco_clean2$nstage=="N0" ~ "Intermediate",
  plco_clean2$nstage=="N0" & (plco_clean2$tstage=="T3a-T4" | plco_clean2$gleason=="8" | plco_clean2$gleason=="9" | plco_clean2$gleason=="10" | plco_clean2$dx_psa > 20) ~ "High",
  plco_clean2$nstage=="N1" | plco_clean2$nstage=="N2" ~ "Regional"
)

plco_clean2$pc_risk_group <- factor(plco_clean2$pc_risk_group, levels=c("Low", "Intermediate", "High", "Regional"), ordered = TRUE)

plco_clean2$primary_tx <- factor(plco_clean2$primary_tx, levels=c("No known treatment", "Other ablative treatment", "Hormone alone", "Radiation alone", "Prostatectomy", "Radiation + hormone"), labels=c("Active surveillance", "Other ablative treatment", "Hormone alone", "Radiation alone", "Prostatectomy", "Radiation + hormone"), ordered=TRUE)


var_label(plco_clean2$age) <- "Age (years)"
var_label(plco_clean2$race2) <- "Race"
var_label(plco_clean2$educ) <- "Education"
var_label(plco_clean2$marital2) <- "Marital status"
var_label(plco_clean2$mortstat) <- "Outcome"
var_label(plco_clean2$permth_exm) <- "Follow-up (months)"
var_label(plco_clean2$smoker) <- "Smoking status"
var_label(plco_clean2$arthritis) <- "Arthritis"
var_label(plco_clean2$bronch) <- "Chronic bronchitis"
var_label(plco_clean2$diabetic) <- "Diabetes"
var_label(plco_clean2$emphysema) <- "Emphysema"
var_label(plco_clean2$hypertension) <- "Hypertension"
var_label(plco_clean2$mi_chd) <- "Previous heart attack, coronary heart disease"
var_label(plco_clean2$liver) <- "Liver disease"
var_label(plco_clean2$stroke) <- "Previous stroke"
var_label(plco_clean2$weight) <- "BMI category"
var_label(plco_clean2$nstage) <- "N Stage"
var_label(plco_clean2$tstage) <- "T Stage"
var_label(plco_clean2$gleason) <- "Gleason"
var_label(plco_clean2$dx_psa) <- "PSA at diagnosis"
var_label(plco_clean2$primary_tx) <- "Treatment"
var_label(plco_clean2$pc_risk_group) <- "Prostate cancer risk group"

table_example <- CreateTableOne(vars = c("age", "race2", "educ", "marital2", "smoker", "diabetic", "hypertension", "stroke", "weight", "nstage"), strata = "causespecific_group", includeNA = TRUE, data = plco_clean2, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

kable(print(table_example, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:12, 14:16, 18:20, 25:28, 30:33))
```

\newpage

eTable 3, cont.: Characteristics of a cohort of 7,596 men from the Prostate, Lung, Colorectal, and Ovarian Cancer Screening Trial, grouped by OCCAM-predicted median survival time. 
```{r, fig.align="center", fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo=FALSE}
table_example2 <- CreateTableOne(vars = c("tstage", "gleason", "dx_psa", "pc_risk_group", "primary_tx", "mortstat"), strata = "causespecific_group", includeNA = TRUE, data = plco_clean2, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

kable(print(table_example2, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(3:7, 9:13, 16:20, 22:27, 29:31))
```
```

